maze-dataset

### 2025-03-30 docs/paper/paper.tex
wip

\hypertarget{training}{%
\subsection{Visual Outputs for Training and Evaluation}\label{training}}

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}). To facilitate the use of our package in this
context, we replicate the format of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) and provide the \texttt{RasterizedMazeDataset}
class which returns rasterized pairs of (input, target) mazes as shown
in \autoref{fig:e2hraster} below.

\begin{figure}
\hypertarget{fig:e2hraster}{%
\centering
\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
\caption{Input is the rasterized maze without the path marked (left),
and provide as a target the maze with all but the correct path removed.
Configuration options exist to adjust whether endpoints are included and
if empty cells should be filled in.}\label{fig:e2hraster}
}
\end{figure}

\hypertarget{tokenized-output-formats}{%
\section{Tokenized Output Formats}\label{tokenized-output-formats}}

@@

\hypertarget{mtm}{%
\subsection{\texorpdfstring{Stringification Options and
\texttt{MazeTokenizerModular}}{Stringification Options and MazeTokenizerModular}}\label{mtm}}

There are many algorithms by which one might tokenize a 2D maze into a
1D format usable by autoregressive text models. Training multiple models
on the encodings output from each of these algorithms may produce very
different internal representations, learned solution algorithms, and
levels of performance. To explore how different maze tokenization
algorithms affect these models, the \texttt{MazeTokenizerModular} class
contains a rich set of options to customize how mazes are stringified.
This class contains 19 discrete parameters, resulting in 5.9 million
unique tokenizers. But wait, there's more! There are 6 additional
parameters available in the library which are untested but further
expand the the number of tokenizers by a factor of \(44/3\) to 86
million.

All output sequences consist of four token regions representing
different features of the maze. These regions are distinguished by color
in Figure below.

@@

\begin{figure}
\centering
\includegraphics{figures/outputs-tokens-colored.tex}
\caption{Example text output format with token regions highlighted.}
\end{figure}

>>>>

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}). To facilitate the use of our package in this
context, we replicate the format of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) and provide the \texttt{RasterizedMazeDataset}
class which returns rasterized pairs of (input, target) mazes as shown
in \autoref{fig:e2h-raster} below.

\begin{figure}
\hypertarget{fig:e2h-raster}{%
\centering
\includegraphics[width=0.3\textwidth,height=\textheight]{figures/maze-raster-input-target.pdf}
\caption{Input is the rasterized maze without the path marked (left),
and provide as a target the maze with all but the correct path removed.
Configuration options exist to adjust whether endpoints are included and
if empty cells should be filled in.}\label{fig:e2h-raster}
}
\end{figure}

\hypertarget{tokenized-output-formats}{%
\subsection{Tokenized Output Formats}\label{tokenized-output-formats}}

@@

There are many algorithms by which one might tokenize a 2D maze into a
1D format usable by autoregressive text models. Training multiple models
on the encodings output from each of these algorithms may produce very
different internal representations, learned solution algorithms, and
levels of performance. To allow exploration of how different maze
tokenization algorithms affect these models, the
\texttt{MazeTokenizerModular} class contains a rich set of options to
customize how mazes are stringified. This class contains 19 discrete
parameters, resulting in over 5.8 million unique tokenizers. But wait,
there's more! There are 6 additional parameters available in the library
which are untested but further expand the the number of tokenizers by a
factor of \(44/3\) to 86 million.

All output sequences consist of four token regions representing
different features of the maze. These regions are distinguished by color
in \autoref{fig:token-regions}.

@@

\begin{figure}
\hypertarget{fig:token-regions}{%
\centering
\includegraphics{figures/outputs-tokens-colored.tex}
\caption{Example text output format with token regions
highlighted.}\label{fig:token-regions}
}
\end{figure}
